Incremental λ-Calculus in Cache-Transfer Style

### 2017-07-10 popl18/soundness.tex
cts-case-studies: abstract over benchmark file location
Stop line breaks at start of theorems
These line breaks don't make sense here, since theorems already start with a
line break (so we get two).

\input{differentiation-and-static-caching-continued}

@@

\begin{lemma}[Evaluation commutes with CTS translation of base terms]
  \-\\
  For all $\sectx, \sterm$ and $\sclosedvalue$, such that
  $\seval\sectx\sterm{}\sclosedvalue$, and
  for all $\tcache$, there exists $\tcachedclosedvalues$,
  $\seval
  {\compile{\sectx}}{\compileterm\tcache\sterm}
  {}
  {(\compile{\sclosedvalue}, \tcachedclosedvalues)}$.
\end{lemma}

@@

\begin{lemma}[Evaluation commutes with compilation of derivatives]
  \-\\
  Let $\tcache$ be such that
  $(\tcache, \_) = \compileterm{\bullet}{\sterm}$.
  For all $\iectx, \sterm$ and $\icloseddvalue$,
  if
  $\seval{\iectx}{\iderive\sterm}{}{\icloseddvalue}$,
  and $\envwithcache{\compile\iectx}{\tcache}{\tdectx}$,
  then
  $\seval{\tdectx}{\derive{\tupdcache}{\sterm}}{}{(\compile\icloseddvalue, \tcachedclosedvalues)}$.
\end{lemma}

@@

\begin{theorem}[Soundness of differentiation with static caching wrt differentiation]
  \-\\
  \label{thm:soundness-compiled-changesfinal}
  Let $\tcache$ and $\tterm$ be such that 
  $(\tcache, \tterm) = \compileterm{\bullet}{\sterm}$.
  For all $\iectx, \sterm$ and $\icloseddvalue$,
  $\seval{\iectx}{\iderive\sterm}{}{\icloseddvalue}$,
  for $\tectx$ and $\tectx'$ such that
  $\envwithcache{\compile\iectx}{\tcache}{\tectx}$
  and $\newenvwithcache{\compile\iectx}{\tcache}{\tectx'}$,
  $\seval{\oldenv\tectx}{\tterm}{}{(\tclosedvalue, \tcachedclosedvalues)}$ and
  $\seval{\newenv{\tectx'}}{\tterm}{}{(\tclosedvalue', \tcachedclosedvalues')}$,
  then
  $
  \seval{\tectx}{\derive\tcache\sterm}{}{(\compile{\icloseddvalue}, \tcachedclosedvalues')}
  $.
\end{theorem}

>>>>

\input{\poplPath{differentiation-and-static-caching-continued}}

@@

\begin{lemma}[Evaluation commutes with CTS translation of base terms]
  For all $\sectx, \sterm$ and $\sclosedvalue$, such that
  $\seval\sectx\sterm{}\sclosedvalue$, and
  for all $\tcache$, there exists $\tcachedclosedvalues$,
  $\seval
  {\compile{\sectx}}{\compileterm\tcache\sterm}
  {}
  {(\compile{\sclosedvalue}, \tcachedclosedvalues)}$.
\end{lemma}

@@

\begin{lemma}[Evaluation commutes with compilation of derivatives]
  Let $\tcache$ be such that
  $(\tcache, \_) = \compileterm{\bullet}{\sterm}$.
  For all $\iectx, \sterm$ and $\icloseddvalue$,
  if
  $\seval{\iectx}{\iderive\sterm}{}{\icloseddvalue}$,
  and $\envwithcache{\compile\iectx}{\tcache}{\tdectx}$,
  then
  $\seval{\tdectx}{\derive{\tupdcache}{\sterm}}{}{(\compile\icloseddvalue, \tcachedclosedvalues)}$.
\end{lemma}

@@

\begin{theorem}[Soundness of differentiation with static caching wrt differentiation]
  \label{thm:soundness-compiled-changesfinal}
  Let $\tcache$ and $\tterm$ be such that 
  $(\tcache, \tterm) = \compileterm{\bullet}{\sterm}$.
  For all $\iectx, \sterm$ and $\icloseddvalue$,
  $\seval{\iectx}{\iderive\sterm}{}{\icloseddvalue}$,
  for $\tectx$ and $\tectx'$ such that
  $\envwithcache{\compile\iectx}{\tcache}{\tectx}$
  and $\newenvwithcache{\compile\iectx}{\tcache}{\tectx'}$,
  $\seval{\oldenv\tectx}{\tterm}{}{(\tclosedvalue, \tcachedclosedvalues)}$ and
  $\seval{\newenv{\tectx'}}{\tterm}{}{(\tclosedvalue', \tcachedclosedvalues')}$,
  then
  $
  \seval{\tectx}{\derive\tcache\sterm}{}{(\compile{\icloseddvalue}, \tcachedclosedvalues')}
  $.
\end{theorem}

### 2017-07-10 popl18/static-differentiation-soundness.tex
cts-case-studies: abstract over benchmark file location
Stop line breaks at start of theorems
These line breaks don't make sense here, since theorems already start with a
line break (so we get two).

\begin{lemma}[$\oplus$ agrees with validity]
  \label{lemma:crel-oplus}
  \-\\
  If
  $\vvcrel {\sclosedvalue_2} k {\sclosedvalue_1} \icloseddvalue$ then
  $\sclosedvalue_1 \oplus \icloseddvalue = \sclosedvalue_2$.
\end{lemma}
\begin{lemma}[Downward-closure]
  \label{lemma:vvcrel-antimono}
  \-\\
  If $N \ge n$, then $\vvcrel{\sclosedvalue_2} {N} {\sclosedvalue_1} \icloseddvalue$ implies
  $\vvcrel{\sclosedvalue_2} {n} {\sclosedvalue_1} \icloseddvalue$.
\end{lemma}

\input{differentiation-step-indexed-relation}

@@

The above lemmas about validity for values extend to environments.
\begin{lemma}[$\oplus$ agrees with validity, for environments]
  \label{lemma:crel-oplus-env}
  \-\\
  If
  $\vvcrel {\sectx_2} k {\sectx_1} \iectx$ then
  $\sectx_1 \oplus \iectx = \sectx_2$.
\end{lemma}
\begin{lemma}[Downward-closure, for environments]
  \label{lemma:vvcrel-antimono-env}
  \-\\
  If $N \ge n$, then $\vvcrel{\sectx_2} {N} {\sectx_1} \iectx$ implies
  $\vvcrel{\sectx_2} {n} {\sectx_1} \iectx$.
\end{lemma}

@@

We can state a soundness theorem for differentiation without
mentioning step-indexes. Instead of proving it directly, we must first
prove a more technical statement
(Lemma~\ref{lemma:fundamental-property}) that mentions step-indexes
explicitly.
%
\begin{theorem}[Soundness of differentiation in $\source$]
  \label{thm:sound-derive}
  \-\\
  If $\iectx$ is a valid change environment from base environment $\sectx_1$ to
  updated environment $\sectx_2$, that is $\vvcrel {\sectx_2} {} {\sectx_1}\iectx$,
  %
  and if $\sterm$ converges both in the base and updated environment, that is
  $\sneval{\sectx_1}\sterm{\sclosedvalue_1}$ and
  $\sneval{\sectx_2}\sterm{\sclosedvalue_2}$,
  %
  then $\iderive\sterm$ evaluates under the change environment~$\iectx$ to a valid
  change $\icloseddvalue$ between base result ${\sclosedvalue_1}$ and updated result
  ${\sclosedvalue_2}$, that is
  $\sneval\iectx{\iderive\sterm}{\icloseddvalue}$ and
  $\vvcrel{\sclosedvalue_2} {} {\sclosedvalue_1}
  \icloseddvalue$.
\end{theorem}

\begin{lemma}[Fundamental Property]
  \label{lemma:fundamental-property}
  \-\\
For each $n$,
if $\vvcrel{\sectx_2}{n}{\sectx_1}{\iectx}$ then
$\vvcrelterm{\sectx_1}{\iectx}{\sectx_2}{n}{\sterm}{\iderive{\sterm}}{\sterm}$.
\end{lemma}

>>>>

\begin{lemma}[$\oplus$ agrees with validity]
  \label{lemma:crel-oplus}
  If
  $\vvcrel {\sclosedvalue_2} k {\sclosedvalue_1} \icloseddvalue$ then
  $\sclosedvalue_1 \oplus \icloseddvalue = \sclosedvalue_2$.
\end{lemma}
\begin{lemma}[Downward-closure]
  \label{lemma:vvcrel-antimono}
  If $N \ge n$, then $\vvcrel{\sclosedvalue_2} {N} {\sclosedvalue_1} \icloseddvalue$ implies
  $\vvcrel{\sclosedvalue_2} {n} {\sclosedvalue_1} \icloseddvalue$.
\end{lemma}

\input{\poplPath{differentiation-step-indexed-relation}}

@@

The above lemmas about validity for values extend to environments.
\begin{lemma}[$\oplus$ agrees with validity, for environments]
  \label{lemma:crel-oplus-env}
  If
  $\vvcrel {\sectx_2} k {\sectx_1} \iectx$ then
  $\sectx_1 \oplus \iectx = \sectx_2$.
\end{lemma}
\begin{lemma}[Downward-closure, for environments]
  \label{lemma:vvcrel-antimono-env}
  If $N \ge n$, then $\vvcrel{\sectx_2} {N} {\sectx_1} \iectx$ implies
  $\vvcrel{\sectx_2} {n} {\sectx_1} \iectx$.
\end{lemma}

@@

We can state a soundness theorem for differentiation without
mentioning step-indexes. Instead of proving it directly, we must first
prove a more technical statement
(Lemma~\ref{lemma:fundamental-property}) that mentions step-indexes
explicitly.
%
\begin{theorem}[Soundness of differentiation in $\source$]
  \label{thm:sound-derive}
  If $\iectx$ is a valid change environment from base environment $\sectx_1$ to
  updated environment $\sectx_2$, that is $\vvcrel {\sectx_2} {} {\sectx_1}\iectx$,
  %
  and if $\sterm$ converges both in the base and updated environment, that is
  $\sneval{\sectx_1}\sterm{\sclosedvalue_1}$ and
  $\sneval{\sectx_2}\sterm{\sclosedvalue_2}$,
  %
  then $\iderive\sterm$ evaluates under the change environment~$\iectx$ to a valid
  change $\icloseddvalue$ between base result ${\sclosedvalue_1}$ and updated result
  ${\sclosedvalue_2}$, that is
  $\sneval\iectx{\iderive\sterm}{\icloseddvalue}$ and
  $\vvcrel{\sclosedvalue_2} {} {\sclosedvalue_1}
  \icloseddvalue$.
\end{theorem}

\begin{lemma}[Fundamental Property]
  \label{lemma:fundamental-property}
For each $n$,
if $\vvcrel{\sectx_2}{n}{\sectx_1}{\iectx}$ then
$\vvcrelterm{\sectx_1}{\iectx}{\sectx_2}{n}{\sterm}{\iderive{\sterm}}{\sterm}$.
\end{lemma}
